Viscous Erosion in Porous Media

### 2018-03-29 docs/methods/erosion_methods.tex
NM

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For a {\em circular} cylinder, the normalization factor $4 \pi$ produces an {\em exact} relationship (within the context of slender-body theory), where $a$ is taken to be the radius. For other shapes the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale since it generalizes the case of a circle. In Fig.~\ref{area_drag}(b) we show this prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ is the radius of the initially circular body. The scaling law and measured drag agree closely as $t \to t_f$, confirming that the simulation accurately captures the slender limit as the body disappears.

@@

	We show in Fig.~\ref{fixed_intface}(a) several interfaces of this modified erosion process, with the same spatial resolution, $N_\iin = 1024$, as before. Notice the body obtains a similar profile as observed previously, but with fixed area. Importantly, more numerical accuracy is now available to quantify the terminal state. The inset in Fig.~\ref{fixed_intface}(a) zooms on the rear corner demonstrate that, due to our numerical regularization, the body is actually smooth on a fine scale. We show in Fig.~\ref{fixed_intface}(b) the stress distributions on the same interfaces. As observed previously, the stress becomes more evenly distributed as erosion reshapes the body. The slight rise in shear near the front/rear corners that was observed in Fig.~\ref{shrink_intface}(b) is no longer present, suggesting that this effect was due to loss of accuracy. Rather, the late-time stress distributions in Fig.~\ref{fixed_intface}(b) decrease monotonically to zero near the stagnation points. In this figure, time is normalized by the vanishing time $t_f = 1.79 \times 10^{-2}$ from the shrinking simulations (i.e.~Fig.~\ref{shrink_intface}). Curiously, about 3 multiples of $t_f$ are required to reach a visible steady state, which contrasts with previous studies on high-Reynolds-number erosion in which a steady state usually emerges before the body vanishes \cite{moo-ris-chi-zha-she2013}.

@@

	These fixed-area simulations enable us to precisely characterize the changing geometry as it approaches a terminal state. In particular, we measure the aspect ratio of the body (the width to height) and the opening angle of the front/rear corners. In these tests, we keep the spatial resolution fixed, $N_\iin = 1024$, and show three different cases of the smoothing parameters to assess how the numerical regularization influences shape. In all three cases, we keep the two parameters equal $\eps = \sigma$ and set to the values 20/1024, 10/1024, and 5/1024 respectively. Figure~\ref{fig:arangle}(a) shows the aspect ratio as it varies in time. The initially circular geometry has an aspect ratio of unity, and, as erosion carves a more slender form, the aspect ratio increases monotonically. The three simulations are all very close to one another, demonstrating convergence as $\eps, \sigma \to 0$. In the simulation with these least smoothing ($\eps = \sigma = 5/1024$), the aspect ratio tends to a value of 2.7 in long time.

	Next, we aim to measure the opening angle of the front/rear corner. Since these simulated bodies are actually smooth on a fine scale, we must have a procedure to estimate the apparent angle that results from the high (but finite) curvature at the stagnation points. To this end, we fit the tangent angle $\theta(\alpha)$ with a 7th-degree polynomial over a region excluding the stagnation points ($\alpha = 0$ and 0.5). We then extrapolate the polynomial to the stagnation points, which gives an estimate of the angle there. Although this procedure is subject to some uncertainty (to be examined below), we have found it to give robust angle measurements. Figure \ref{fig:arangle}(b) shows the opening angle, $\beta$, as it varies in time for the three cases of smoothing parameters. For the initially circular geometry, our procedure yields the correct result of a $180^{\circ}$ opening angle, i.e.~a flat front face. As time proceeds, erosion carves a sharper corner and $\beta$ decreases monotonically. Here, the three curves are more easily distinguished than in Fig.~\ref{fig:arangle}(a), implying that the opening angle is more sensitive to the level of numerical smoothing. Nonetheless, the curves show convergence as $\eps, \sigma \to 0$. In the finest simulation ($\eps = \sigma = 5/1024$), $\beta$ converges to the value $102^{\circ}$, which agrees precisely with the analytical prediction from Section \ref{LimitingShape}.

	To more thoroughly examine the influence of the smoothing parameters, we report in Table~\ref{table:arangle} the long-time values of the aspect ratio and opening-angle measurements for the three cases. Loosely, with $N_\iin = 1024$ fixed, the values 20/1024, 10/1024, and 5/1024 correspond to smoothing over the nearest 20, 10, and 5 grid points respectively. As seen in the table, both the aspect ratio and the opening angle converge as $\eps, \sigma \to 0$ with roughly first order. To quantify the uncertainty involved in our angle measurement procedure, we show the estimated uncertainty, which is obtained by varying the degree of the fitting polynomial and the size of the excluded region around the stagnation point....

>>>>

As a check on these results, we can use slender-body theory to estimate the drag-per-unit-span
\begin{align}
\label{dragscaling}
\frac{F_D}{4 \pi} \sim \frac{ \mu U}{\log(\ell/a)}	\qquad \text{for } a \ll \ell \, .
\end{align}
For a {\em circular} cylinder, the normalization factor $4 \pi$ produces an {\em exact} relationship (within the context of slender-body theory), where $a$ is taken to be the radius. For other shapes, the formula will be approximate and we define $a = \sqrt{A/\pi}$ as the characteristic lengthscale since it generalizes the case of a circle. In Fig.~\ref{area_drag}(b) we show this prediction (red dotted curve), where $\ell = 2.3 a_0$ is a fit parameter and $a_0 = 0.2$ is the radius of the initially circular body. The scaling law and measured drag agree closely as $t \to t_f$, confirming that the simulation accurately captures the slender limit as the body disappears.

@@

	We show in Fig.~\ref{fixed_intface}(a) several interfaces of this modified erosion process, with the same spatial resolution, $N_\iin = 1024$, as before. Notice the body obtains a similar profile as observed previously, but now with fixed area. Importantly, more numerical accuracy is available to quantify the terminal state. The inset in Fig.~\ref{fixed_intface}(a) zooms on the rear corner demonstrate that, due to our numerical regularization, the body is actually smooth on a fine scale. We show in Fig.~\ref{fixed_intface}(b) the stress distributions on the same interfaces. As observed previously, the stress becomes more evenly distributed as erosion reshapes the body. The slight rise in shear near the front/rear corners that was observed in Fig.~\ref{shrink_intface}(b) is no longer present, suggesting that this effect was due to loss of accuracy. Rather, the late-time stress distributions in Fig.~\ref{fixed_intface}(b) decrease monotonically to zero near the stagnation points. In this figure, time is normalized by the vanishing time $t_f = 1.79 \times 10^{-2}$ from the shrinking simulations (i.e.~Fig.~\ref{shrink_intface}). Curiously, about 3 multiples of $t_f$ are required to reach a visible steady state, which contrasts with previous studies on high-Reynolds-number erosion in which a steady state usually emerges before the body vanishes \cite{moo-ris-chi-zha-she2013}.

@@

	These fixed-area simulations enable us to precisely characterize the changing geometry as it approaches the terminal state. In particular, we measure the aspect ratio of the body (the width to height) and the opening angle of the front/rear corners. In these tests, we keep the spatial resolution fixed, $N_\iin = 1024$, and show three different cases of the smoothing parameters, $\eps$ and $\sigma$, to assess how the numerical regularization influences shape. In the three cases, we keep $\eps = \sigma$ and set them to the values 20/1024, 10/1024, and 5/1024 respectively. Figure~\ref{fig:arangle}(a) shows the aspect ratio as it varies in time. The initially circular geometry has an aspect ratio of unity, and, as erosion carves a more slender form, the aspect ratio increases monotonically. The three simulations are all very close to one another, demonstrating convergence as $\eps, \sigma \to 0$. In the simulation with these least smoothing ($\eps = \sigma = 5/1024$), the aspect ratio tends to a value of 2.7 in long time.

	Next, we aim to measure the opening angle of the front/rear corner. Since these simulated bodies are actually smooth on a fine scale, we must have a procedure to estimate the apparent angle that results from the high (but finite) curvature at the stagnation points. To this end, we fit the tangent angle $\theta(\alpha)$ with a 7th-degree polynomial over a region excluding the stagnation points ($\alpha = 0$ and 0.5). We then extrapolate the polynomial to the stagnation points, which gives an estimate of the angle there. Although this procedure is subject to some uncertainty (to be examined below), we have found it to give robust angle measurements. Figure \ref{fig:arangle}(b) shows the opening angle, $\beta$, as it varies in time for the three cases of smoothing parameters. For the initially circular geometry, our procedure yields the correct opening angle of $180^{\circ}$, i.e.~a flat front face. As time proceeds, erosion carves a sharper corner and $\beta$ decreases monotonically. Here, the three simulation curves are more easily distinguished, implying that the opening angle is more sensitive to the level of numerical smoothing. The curves nevertheless show convergence as $\eps, \sigma \to 0$. In the finest simulation ($\eps = \sigma = 5/1024$), $\beta$ converges to the value $102^{\circ}$, which agrees precisely with the analytical prediction from Section \ref{LimitingShape}.

	To more thoroughly examine the influence of the smoothing parameters, we report in Table~\ref{table:arangle} the long-time values of the aspect ratio and opening angle for each of the three simulations. Loosely, with $N_\iin = 1024$ fixed, the values 20/1024, 10/1024, and 5/1024 correspond to smoothing over the nearest 20, 10, and 5 grid points respectively. As seen in the table, both the aspect ratio and the opening angle converge as $\eps, \sigma \to 0$ with roughly first order. To quantify the uncertainty involved in our angle measurement procedure, we show the estimated uncertainty, which is obtained by varying the degree of the fitting polynomial and the size of the excluded region around the stagnation point....
